---
title: "Master Thesis 4.0 Misc"
author: "Marijan Rancic"
date: "10/10/2020"
output: html_document
editor_options: 
chunk_output_type: inline
---

## Set working directory
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE, message=FALSE)
knitr::opts_knit$set(root.dir = "/Users/marijanrancic/Documents/IMQF/Master/QuantRSpace/MasterProject/FinalMasterWork/")
getwd()
```
## Note on R format
Main version of the thesis will be delivered in RM. 
This file has been initiated and changed in the Final Version as a file subsequent to PredictiveModeling_Master

```{r libraries, warning=FALSE, include=FALSE}
# Think about only necessary libraries because of the masking and if you've already initiated Predictive Modeling this is not mandatory

library(readxl)
library(moments)
library(tseries)
# library(MASS)
library(stats)
library(ggplot2)
library(tidyverse)
library(fpp3)
#library(forecast)
#library(reshape2)
# Depends on the file that are initiated
```

This part will be to create some data frames and clean/arrange various instances of SEEPEX and HUPX DATA

## Preparatory data work

Data has been provided as Excel document. Make sure that you have data in MS SQL database at some point of time. 

We decided to delete two observation with - prices on HUPX and few with zero price and replace them with price of 0.1EUR
```{r}
DBTS %>% filter(RS<=0)
DBTS %>% filter(HU<=0)
```

```{r BasicCleaning}
DBTS$HU[which(DBTS$HU <=0)] = 0.1
DBTS$RS[which(DBTS$RS <=0)] = 0.1
```
And the results:
```{r}
DBTS %>% filter(RS<=0)
DBTS %>% filter(HU<=0)
```


```{r Data_Filter, echo=True}

# Provera da li imamo NA vrednosti cena u datasetu.
# HUPX NA
idx1 <- which(is.na(DBTS$LoadF))
DBTS[idx1,]
idx2 <- which(is.na(DBTS$temp))
DBTS[idx1,]

```

#Useful for further work
If want to create iterable lists we can use this below. This has been developed in a separate file. It is great feature for the com.

```{r}
iterable.DBTS <- DBTS %>% 
mutate(hour_num = as.factor(hour_num)) %>% group_by(hour_num) %>% 
  group_split() %>% 
map_dfr(.f = function(df){
 lm(RS ~ HU, data=df) %>% 
    glance() %>% 
    add_column(test = unique(df$hour_num),.before = 1)
}
  )
```

# Basic Data Graphics
## Price 

```{r echo=TRUE}
library(ggplot2)
library(gridExtra)

g1 <- ggplot(DBTS, aes(x=date.time)) + 
  geom_line(aes(y=HU)) + 
  labs(  
       subtitle="HUPX Spot Price", 
       x="Date",
       y="EUR/MWh")
ggsave(file="Output/DBTS-HU.jpg",g1,width = 18, height = 12, dpi = 300, units = "cm")

# Second graph
coeff <- 1 #scale levog i desnog grafika, ako je 10 na desnom ce biti 10puta veca vrednost
g2 <- ggplot(DBTS %>% filter(year(date.time)==2017 | year(date.time)==2018), aes(x=date.time)) +
  geom_line( aes(y=HU), color='blue') + 
  geom_line( aes(y=RS), color='brown') + 
  scale_y_continuous(
    name = "Hupx peak price - brown", 
    sec.axis = sec_axis(~.*coeff, name="Seepex peak price - blue")
  ) + 
  labs(  caption="Source: Author's own", 
       subtitle="HUPX-SEEPEX Spread in-sample period")

ggsave(file="Output/DBTS-HU-RS Spread.jpg",g2,width = 18, height = 12, dpi = 300, units = "cm")

```



Two ways to present Hourly prices on graph. 
However, first one would generate error if myfunctions are not imported and column hour is erased. 

```{r}
library(ggplot2)

# ggplot(DBHour("H15"), aes(x=date.time)) + 
#   geom_line(aes(y=HU)) + 
#   labs(
#        subtitle="HUPX Hourly price", 
#        x="Date",
#        y="EUR/MWh")

ggplot(DBTS %>% filter(DHour_H15==1), aes(x=datum)) + 
  geom_line(aes(y=HU)) + 
  labs(
       subtitle="HUPX price ar predifined hour", 
       x="Date",
       y="EUR/MWh")
```
```{r}

g1 <- ggplot(DBTS %>% filter(year(date.time)==2017 & quarter(date.time)==1), aes(x=date.time)) + 
  geom_line(aes(y=RS)) + 
  labs(subtitle = "SEEPEX Spot Price, Q1 2017",
       x="Date",
       y="EUR/MWh")


ggsave(file="Output/spikes_seepex2017_q1.jpg",g1,width = 18, height = 12, dpi = 300, units = "cm")
```

```{r}
library(gridExtra)
g1 <- ggplot(DBTS %>% filter(year(date.time)==2018 & hour(date.time)>=16 & hour(date.time)<20 & quarter(date.time)==1), aes(x=date.time)) + 
  geom_line(aes(y=RS, colour=DHour)) + 
  labs(subtitle = "SEEPEX (17-20) Rush hour, Q1 2018",
    colour ="Hour", 
       x="Date",
       y="EUR/MWh")
g2 <- ggplot(DBTS %>% filter(year(date.time)==2018 & hour(date.time)>=00 & hour(date.time)<04 & quarter(date.time)==1), aes(x=date.time)) + 
  geom_line(aes(y=RS, colour=DHour)) + 
  labs(subtitle = "SEEPEX (00-04) Off Peak, Q1 2018",
       colour="Hour",
      caption="Source: Author's own", 
       x="Date",
       y="EUR/MWh")

gfinal <- grid.arrange (g1,g2)
ggsave(file="Output/price_rushhours_quat20181.jpg",gfinal)
rm(g1,g2)
```


```{r}
library(gridExtra)


p1 <- ggplot(DBTS %>% filter(date(date.time)=="2017-04-14" | date(date.time)=="2017-04-15" | date(date.time)=="2017-04-16"| date(date.time)=="2017-04-17"), aes(x=date.time)) + 
  geom_line(aes(y=HU)) + 
  labs(
      subtitle ="2017", 
       x="Date",
       y="EUR/MWh")

p2 <- ggplot(DBTS %>% filter(date(date.time)=="2018-04-14" | date(date.time)=="2018-04-15" | date(date.time)=="2018-04-16"| date(date.time)=="2018-04-17" ), aes(x=date.time)) + 
  geom_line(aes(y=HU)) + 
  labs(
      subtitle ="2018", 
       x="Date",
       y="EUR/MWh")
p3 <- ggplot(DBTS %>% filter(date(date.time)=="2019-04-14" | date(date.time)=="2019-04-15" | date(date.time)=="2019-04-16"| date(date.time)=="2019-04-17"), aes(x=date.time)) + 
  geom_line(aes(y=HU)) + 
  labs(caption="Source: Author's own",
      subtitle ="2019", 
       x="Date",
       y="EUR/MWh")

pfinal <- grid.arrange (p1,p2,p3)
ggsave(file="Output/4daysSEEPEX.jpg",pfinal)


```
```{r}
earlymorning <- DBTS %>% filter(year(date.time)==2018 & hour(date.time)>=8 & hour(date.time)<11) %>% summarise(meanprice=mean(HU))
earlymorning %>% autoplot(meanprice) +
  ylab("Mean early morning price") + xlab("Year") 
```



## Returns and statistic of FRM data
Here we will process FRM data

```{r echo=TRUE}
library(ggplot2)

#izbacicu ove outliere
frm_smooth1 <-  daily_prices %>% 
  filter(peak_price <(mean(peak_price) + 3.3 * sd(peak_price)))

ggplot(data_frm, aes(x=date)) + 
  geom_line(aes(y=peak_price)) + 
  labs(
       subtitle="SUPX price price 8AM-8PM", 
       x="Date",
       y="EUR/MWh")

```

## Histogram

```{r Graph_HistogramANDPRicesperHOUR}
library(ggplot2)
theme_set(theme_bw())
p1 <- ggplot(DBHour("H04"), aes(x=date.time)) + 
  geom_line(aes(y=HU)) + 
  labs(
       subtitle="Hour 4", 
       x="Date",
       y="EUR/MWh")

p2 <- ggplot(DBHour("H04"), aes(HU)) +
  geom_histogram(bins = 100,
                 col = "grey95",
                 fill = "black", 
                 alpha = .80) +
  labs(x = "EUR/MWh", y = "Frequency",subtitle="Hour 4") +
  theme(title = element_text(size = 12),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10, face = "bold"))
p3 <- ggplot(DBHour("H10"), aes(x=date.time)) + 
  geom_line(aes(y=HU)) + 
  labs(
       subtitle="Hour 10", 
       x="Date",
       y="EUR/MWh")

p4 <- ggplot(DBHour("H10"), aes(HU)) +
  geom_histogram(bins = 100,
                 col = "grey95",
                 fill = "black", 
                 alpha = .80) +
  labs(x = "EUR/MWh", y = "Frequency",subtitle="Hour 10") +
  theme(title = element_text(size = 12),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10, face = "bold"))

p5 <- ggplot(DBHour("H20"), aes(x=date.time)) + 
  geom_line(aes(y=HU)) + 
  labs(
       subtitle="Hour 20", 
       x="Date",
       y="EUR/MWh")

p6 <- ggplot(DBHour("H20"), aes(HU)) +
  geom_histogram(bins = 100,
                 col = "grey95",
                 fill = "black", 
                 alpha = .80) +
  labs(x = "EUR/MWh", y = "Frequency",subtitle="Hour 20") +
  theme(title = element_text(size = 12),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10, face = "bold"))
pfinal <- grid.arrange (p1,p2,p3,p4,p5,p6,ncol=2,
                layout_matrix = rbind(c(1, 2),
                        c(3, 4),c(5,6))
)
ggsave(file="/Users/marijanrancic/Documents/IMQF/Master/Master teza TEXT/Slike/DBTS-HU-3x2Hourly2.jpg",pfinal,width = 18, height = 12, dpi = 300, units = "cm")
rm(p1,p2,p3,p4,p5,p6,pfinal)
```

## Scatter plot
```{r Graph_Scatter}
library(ggplot2)

theme_set(theme_bw())  # pre-set the bw theme.

# g <- ggplot(DBTS, aes(x=temp,y=LoadF,fill=DHour))
# # Scatter plot
# g + geom_point() + 
#   geom_smooth(method="lm", se=F) +
#   labs(y="Load Forecast", 
#        x="Temperature")


g <- ggplot(DBTS %>% filter(year(date.time)==2018 & hour(date.time)==06), aes(x=TempHU,y=FLoadHU,shape=DHour)) 
g + geom_point() + 
geom_smooth(method="lm", se=F) +
labs(y="Load Forecast", 
x="Temperature")

```
##Napomena
Ovaj grafik nece dati previse neku korelaciju. Trebalo bi videti samo sta se desava u zimskim mesecima, a to radimo ispod


The next thing we are doing is Meteorological seasons filtering. 


```{r}
library(dplyr)

g <- ggplot(DBTS, aes(HU, RS))
g1 <- g + geom_jitter(aes(col=DHour)) + 
  geom_smooth(aes(col=DHour), method="lm", se=F)+
   labs(
        y="SEEPEX hourly prices", 
        x="HUPX hourly pricese")

ggsave(file="/Users/marijanrancic/Documents/IMQF/Master/Master teza TEXT/Slike/DBTS-HU-RS-HCorrelation.jpg",g1,width = 18, height = 12, dpi = 300, units = "cm")
rm(g1)
g <- ggplot(DBTS.SPREAD, aes(HU, RS,colour=Flag))
g + geom_jitter()+
   labs(
        y="SEEPEX daily prices", 
        x="HUPX daily pricese")
```



```{r Descriptive Statistic}
library(moments)
library(tseries)
hourlyprices <- DBTS$HU
show_moments(hourlyprices)
adf.test(hourlyprices)


#“The cross-correlation function helps you discover lagged correlations between two time series. A lagged correlation occurs when today’s value in one time series #is correlated with a future or past value in the other time series.”

ccf(DBTS$HU,DBTS$FLoadHU, main="Price vs Temp")


 
```


```{r Financial assets vs Electricity Price, echo=FALSE}

 
 #UZECU DJI INDEX RETURNS ZA ISTI PERIOD DA UPOREDIM ---------------
 #------KOMENT JER SAM VEC TO URDIO -
# dji_raw <- read_excel("/Users/marijanrancic/Documents/IMQF/Master/Data/DJI2017-2019.xlsx")
# dji.ts <- dji_raw %>%
#   mutate(datum= ymd(datum)) %>%
#   as_tsibble(index =datum)
# 
# peak <- read_excel("/Users/marijanrancic/Documents/IMQF/Master/Data/FRM/DB_FRM.xlsx", sheet = 1)
# peak.ts <- peak %>%
#   mutate(datum= ymd(datum)) %>%
#   as_tsibble(index =datum)
# 
# ts_com <- left_join(dji.ts , peak.ts ,
#           by = c("datum" = "datum")) %>% filter(!is.na(Close))


# names(ts_com)[4] <- "ReturnsPeakPrice"
# names(ts_com)[5] <- "ReturnsDJI"
#--------------------------------------------------------------------
 library(gridExtra)

# - grafik sa returns
coeff <- 1 #scale levog i desnog grafika, ako je 10 na desnom ce biti 10puta veca vrednost
p3 <- ggplot(ts_com, aes(x=datum)) +
  geom_line( aes(y=SimpleRet), color='red') + 
  geom_line( aes(y=Returns), color='blue') + 
  scale_y_continuous(
    name = "Returns",
    sec.axis = sec_axis(~.*coeff)
  ) + 
  labs(subtitle = "Daily returns on DJI index (red) and HUPX peak price (blue)",
      
       x="Year",
       y="EUR/MWh")
# - grafik sa cenama

p1 <- ggplot(ts_com, aes(x=datum)) +
  geom_line( aes(y=peak_price), color='blue') +labs(y="HUPX peak price", x="Year")
p2 <- ggplot(ts_com, aes(x=datum)) +
  geom_line( aes(y=Close), color='red') +labs(y="DJI Index", x="Year")

pfinal <- grid.arrange (p1,p2,p3,ncol=2,
                layout_matrix = rbind(c(1, 2),
                        c(3, 3))
)

ggsave(file="/Users/marijanrancic/Documents/IMQF/Master/Master teza TEXT/Slike/dji_hupx1.jpg",pfinal)



```

# Learning part 
## Fitting examples form the NET
Learning process that will be used for my master thesis



```{r Learning_Forecast1}

library(fpp3)
y <- DBTS
y1 <- DBTS.SPREAD

```


```{r}


y %>%
  filter(year(date.time) > 2018) %>%
  autoplot(HU) +
    xlab("Year: 2017") + ylab(NULL) +
    ggtitle("Load Forecast")
```
# Trend and seasonality
When data are seasonal, the autocorrelations will be larger for the seasonal lags (at multiples of the seasonal frequency) than for other lags.

```{r}
p1 <- y1 %>% ACF(base.HU, lag_max = 21) %>% autoplot() +xlab("Timelags[days]") + ylab("Autocorrelation") + ggtitle("ACF - HUPX Base price")
p2 <-y1 %>% ACF(ret.art.bHU, lag_max = 21) %>% autoplot()+xlab("Timelags[days]") + ylab("Autocorrelation") + ggtitle("ACF - HUPX Base price returns")
p3 <-y1 %>% PACF(base.HU, lag_max = 21) %>% autoplot()+xlab("Timelags[days]") + ylab("Autocorrelation") + ggtitle("PACF - HUPX Base price")
p4 <-y1 %>% PACF(ret.art.bHU, lag_max = 21) %>% autoplot()+xlab("Timelags[days]") + ylab("Autocorrelation") + ggtitle("PACF - HUPX Base price returns")

pfinal <- grid.arrange(p1,p2,p3,p4)
ggsave(file="/Users/marijanrancic/Documents/IMQF/Master/Data/Images/acf1.jpg",pfinal)
# Playing


```
Ovaj deo je komentiran jer ne daje lepe slike ovde
Moze se iz skripta uraditi


```{r Correlation}
# library(GGally)
# y2 <- y1 %>% 
#   select(c(base_price,generation,loadF,avg_temp)) 
# 
# y2 %>% 
#   GGally::ggpairs(columns = 1:4)
```

#TS components
The output above shows the components of an STL decomposition. The original data is shown (as base_price), followed by the estimated components. This output forms a “dable” or decomposition table. The header to the table shows that the priceseries has been decomposed additively.

Ja sam to dole uzeo kao koment da ne pirkazuje

```{r STL decomosition}
# Bitno je da je y tibble !!!!
#y1 <- y1[-1,]
dcmp <- y1 %>% model(STL(base.HU))
dcmp2 <- y2 %>% model(STL(HU))
#components(dcmp) 
#ovo je comented da be bi prikaivao
y1 %>%
  autoplot(base.HU, color='gray') +
  autolayer(components(dcmp), trend, color='red') +
  xlab("Year") + ylab("HUPX Base Price") 

components(dcmp) %>% autoplot() + xlab("Year")
components(dcmp2) %>% autoplot() + xlab("Year")
y1 %>%
  model(STL(base.HU ~ trend(window=21) + season(window='periodic'),
    robust = TRUE)) %>%
  components() %>%
  autoplot()



```
# Moving Average 
```{r}
#mora prvo y1 da bude ispravan format
mydata1 <- y1 %>%
  mutate(
    mv7 = slider::slide_dbl(base.HU, mean, .before=3, .after=3, .complete=TRUE)
  )
mydata1 %>%
  autoplot(base.HU) +
  autolayer(mydata1, mv7, color='red') +
  xlab("Year") + ylab("Base price") +
  ggtitle("HUPX base price with Moving Average") +
  guides(colour=guide_legend(title="series"))

```

```{r zgodno za master iz knjige}
#Most seasonal hour Mislim da je H8

y1 %>%
  features(spot_price, feat_stl) %>%
  filter(seasonal_strength_week == max(seasonal_strength_week)) %>%
  left_join(y1, by = "hour") %>%
  ggplot(aes(x = date.time, y = spot_price)) + geom_line() +
  facet_grid(vars(hour))

# y1_features <- y %>%
#   features(base_price, feature_set(pkgs="feasts"))
```


# Least square estimation for prices

```{r}
#prvo konvert u TSIBLLE
frm_smooth2 <- frm_smooth1 %>%
  mutate(date= ymd(date)) %>%
  as_tsibble(index =date)

frm_smooth3 <- frm_smooth2 %>%
  mutate(
    mv7 = slider::slide_dbl(peak_price, mean, .before=3, .after=3, .complete=TRUE)
  )
frm_smooth3 <- frm_smooth3 %>% drop_na(mv7) # Remove NA vrednosti

frm_smooth3 %>%
  autoplot(peak_price) +
  autolayer(frm_smooth3 , mv7, color='red') +
  xlab("Year") + ylab("Base price") +
  ggtitle("SEEPEX peak price with Moving Average") +
  guides(colour=guide_legend(title="series"))

# a da probamo STD decompresiju



```


```{r}
DBTS
```

