---
title: "Master Thesis 1.0 - IMPORT DATA"
author: "Marijan Rancic"
date: "10/10/2020"
output: html_document
editor_options: 
chunk_output_type: inline
---


## Set working directory
This Chunk is optional and prior run it, please make sure you have your project working directory set.
In addition, due to numerous versions and files, some code chunks are not the examples of best industry practice but I've had to hardwire some changes.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE, message=FALSE)
knitr::opts_knit$set(root.dir = "/Users/marijanrancic/Documents/IMQF/Master/QuantRSpace/MasterProject/MT_MR/MT_MR/")
getwd()
```

## Problem description
This file serves only for importing data and I will erase importing parts in other files. 

## Load libraries
```{r}
library(readxl)
library(lubridate)
library(fastDummies)
library(tidyverse)
source("myFunc.R") # some functions I've wrote that might help. Some are doubled. Appologize for being a dick!
```

## Import hourly data
```{r}
prices<-read_excel("Data/MainDB.xlsx") #This is under MR working directory structure. Make sure you have the same structure or full link to the files.
holidays<-read_excel("Data/HollidayHU.xlsx")

prices<-prices%>%
  left_join(holidays,by=c("datum"="Date"))%>%
  rename("Holiday"="Flag")%>%
  replace_na(list(Holiday=0))

prices$DHour1<-factor(prices$DHour)
prices$hour_num<-hour(dmy_hm(prices$date.time))
prices$weekday<-factor(weekdays(prices$datum))
prices$month<-factor(months(prices$datum))
prices$HU_RS=prices$HU-prices$RS
```

## Generate dummy variables for weekday, month and hour for hourly data
```{r}
prices <- dummy_cols(prices, 
                     select_columns = c('weekday', 'month','DHour1'),
                     remove_selected_columns = TRUE)
```

## Import daily data
Here I've introduced several redundant datasets. The reason is that I've worked on so many versions that I wanted first to avoid errors. 
Later, I'll start deleting them and streamline the code to only one datasets for particular problem. 
```{r}
# reading data  
daily_prices1=read.csv("Data/MainDB_Daily1.csv") # I needed extra for plotting //has no other purpose
daily_prices<-read_excel("Data/MainDB_daily.xlsx") #This is under MR working directory structure. Make sure you have the same structure or full link to the files.

#Holiday is already included in the list. Otherwise, see above how I've joined table for similar exercise. 
daily_prices$month<-factor(months(daily_prices$datum))
```
## Generate dummy variables for month
```{r}
daily_prices <- dummy_cols(daily_prices, 
                     select_columns = 'month',
                     remove_selected_columns = TRUE)
```


## I've followed Forecasting: Principles and Practice by Rob J Hyndman and George Athanasopoulos
This is again some doubled dataset's created for exploring, graphics, etc. and should be streamlined. 
```{r}
library(fpp3) # it will mask some purr methods so make sure to turn on purr later for particular tasks
DBTS <- prices %>%
  mutate(date.time= dmy_hm(date.time)) %>%
  as_tsibble(index =date.time, key=c(hour_num))
DBTS <- DBTS %>% arrange(., date.time)
```
and daily values as .SPREAD
```{r}
DBTS.SPREAD <- daily_prices %>%
  mutate(datum= ymd(datum)) %>%
  as_tsibble(index =datum)
DBTS.SPREAD <- DBTS.SPREAD %>% arrange(., datum)
```

## Additional set for DCC 
We are using R package rmgarch by Alexios Ghalanos Copyright (C) 2008-2016.
```{r}
y = daily_prices[,c(24,26)]
row.names(y) <- DBTS.SPREAD$datum #Depreciated in some versions
```

