---
title: "Master Thesis 5.1"
author: "Marijan Rancic"
date: "10/11/2020"
output: html_document
editor_options: 
chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE, message=FALSE)
knitr::opts_knit$set(root.dir = "/Users/marijanrancic/Documents/IMQF/Master/QuantRSpace/MasterProject/MT_MR/MT_MR/")
getwd()
```

# LOADING PACKAGES 

```{r}
library(forecast)
library(tseries)
```

# PREPROCESSING 

## Reading data  

```{r}
ls(daily_prices)
```

# The Predicting Variables 

- base.HU
- peak.HU

# The Exogenous Variables 
- WGenRO
- GenHU
- TempHU
- FLoad
- CO2Price

# The dummy variables 

- Days
- Holiday
months


# Stationarity Assesment & Time Series Transformation 

For this, we can use the Augmented Dickey-Fuller (ADF) test and Phillips-Pherron which both tests the null hypothesis that the series is non-stationary. This is included in the "tseries" package.

```{r}
adf.test(daily_prices$base.HU, alternative="stationary")
# adf.test(daily_prices$peak.HU, alternative="stationary")
```

```{r}
PP.test(daily_prices$base.HU)
# PP.test(daily_prices$peak.HU)
```

Hypothesis: 
Null Hypothesis: There is a Unit Root for the series 
Alternative Hypothesis: There is no Unit Root for the series. The series is stationary. 

Interpretaion: OK, p-value is way below 0.05. Looks like our data is indeed stationary. Given 
that we had no reason to difference the data, the d value for our ARIMA model is 0. 


# To illustrate out of sample forecasting with covariates, lets splt the data into train and valid.

```{r}
daily_set=daily_prices[1:1095, ]
daily_set=ts(daily_set, start=1)

```
Here is just one of the instances without months and lags
```{r}
train=daily_set[1:730, ]
valid=daily_set[731:1095, ]
xreg1=as.matrix(train[, c("WGenRO", "GenHU", "FLoadHU", "TempHU", "CO2Price", "d1", "d2", "d3","d4","d5", "d6", "Holiday")])
xreg2=as.matrix(valid[, c("WGenRO", "GenHU", "FLoadHU", "TempHU", "CO2Price", "d1", "d2", "d3","d4","d5", "d6", "Holiday")])
xreg1=xreg1[, -1]
xreg2=xreg2[, -1]
```

## Generate lagged values
We allow our models to include lagged values of the dependent variable (24 hours ago), as well as lagged values of predictors (24 hours ago and 1 hour ago) in addition to variables reflecting weekday, month, hour and the values of predictors in the hour for which the prediction is made.

```{r}
library("Hmisc")
# generate all sensible lags to be used as predictors
daily_prices$base.HU_lag1=Lag(daily_prices$base.HU,1)
daily_prices$peak.HU_lag1=Lag(daily_prices$peak.HU,1)

daily_prices$base.HU_lag7=Lag(daily_prices$base.HU,7)
daily_prices$peak.HU_lag7=Lag(daily_prices$peak.HU,7)

daily_prices$base.RS_lag1=Lag(daily_prices$base.RS,1)
daily_prices$peak.RS_lag1=Lag(daily_prices$peak.RS,1)

daily_prices$base.RS_lag7=Lag(daily_prices$base.RS,7)
daily_prices$peak.RS_lag7=Lag(daily_prices$peak.RS,7)

daily_prices$WGenRO_lag1=Lag(daily_prices$WGenRO,1)
daily_prices$GenHU_lag1=Lag(daily_prices$GenHU,1)
daily_prices$FLoadHU_lag1=Lag(daily_prices$FLoadHU,1)
daily_prices$TempHU_lag1=Lag(daily_prices$TempHU,1)
daily_prices$CO2Price_lag1=Lag(daily_prices$CO2Price,1)

daily_prices$WGenRO_lag7=Lag(daily_prices$WGenRO,7)
daily_prices$GenHU_lag7=Lag(daily_prices$GenHU,7)
daily_prices$FLoadHU_lag7=Lag(daily_prices$FLoadHU,7)
daily_prices$TempHU_lag7=Lag(daily_prices$TempHU,7)


```



# fit model for base price 
Fit and Output Model for Base Price
```{r}

fitModelA=auto.arima(train[, "base.HU"], seasonal = TRUE,allowdrift = TRUE,allowmean = TRUE,xreg=xreg1)
#fitModelB=auto.arima(train[, "HU"], xreg=xreg1)
fitModelB=Arima(train[, "base.HU"], order=c(5,0,0), xreg=xreg1)
summary(fitModelA)
summary(fitModelB)

```





# check the residualities of the model

```{r}
checkresiduals(fitModelA)
checkresiduals(fitModelB)
#checkresiduals(fitModelC)
```

Note: An Auto ARIMAX model was used because the model itself will generate the optimal p, d, and q values which would be suitable for the data set to provide better forecasting. The model parameters are shown above. R estimates the parameters using log likelihood. From the standard errors, we see that all the parameters are more than 2 standard deviations away from zero. So all the parameters passed the t-test. The model also calculated the variance of the error term contained in the model. It also gives the likelihood AIC, and the BIC. 


# To determine whether the errors in our ARIMA forecast are normally distributed with a mean of 0 and constant variance, we can also visualize our forecast errors with a time plot. 

```{r}
plot.ts(fitModelA$residuals)

```

This time plot should show that our forecast errors have generally equal variance over time, and that their mean is around zero

```{r}
fcast1=forecast(fitModelA, xreg=xreg2)
fcast2=forecast(fitModelB, xreg=xreg2)
#fcast3=forecast(fitModelC, xreg=xreg2)
```

```{r}
fcast1

```

Visualisation

```{r}
plot(fcast1)
plot(fcast2)
#plot(fcast3)

```

Forecast Accuracy 

```{r}
accuracy(fcast1)
accuracy(fcast2)

```

##PLOT
This is only plot for manual window of forecast. Make sure that you address properly length on the vector below
```{r}

 library(zoo)
 df <-
   data.frame(real = daily_prices$base.HU[1:942],
              pred = c(rep(NA,length(daily_prices$base.HU[1:942])-length(data.frame(fcast1)[,1])),data.frame(fcast1)[,1]),
              time =  zoo::as.Date(seq(as.Date("2017/1/1"), as.Date("2019/07/31"), by = "day"), format = "%Y-%m-%d"),
              Lo80 =c(rep(NA,length(daily_prices$base.HU[1:942])-length(data.frame(fcast1)[,2])),data.frame(fcast1)[,2]),
              Hi80 =c(rep(NA,length(daily_prices$base.HU[1:942])-length(data.frame(fcast1)[,3])),data.frame(fcast1)[,3]),
              Lo95 =c(rep(NA,length(daily_prices$base.HU[1:942])-length(data.frame(fcast1)[,4])),data.frame(fcast1)[,4]),
              Hi95 =c(rep(NA,length(daily_prices$base.HU[1:942])-length(data.frame(fcast1)[,5])),data.frame(fcast1)[,5])) 
 
 df2 <-
   data.frame(real = daily_prices$base.HU[1:942],
              pred = c(rep(NA,length(daily_prices$base.HU[1:942])-length(data.frame(fcast2)[,1])),data.frame(fcast2)[,1]),
              time =  zoo::as.Date(seq(as.Date("2017/1/1"), as.Date("2019/07/31"), by = "day"), format = "%Y-%m-%d"),
              Lo80 =c(rep(NA,length(daily_prices$base.HU[1:942])-length(data.frame(fcast2)[,2])),data.frame(fcast2)[,2]),
              Hi80 =c(rep(NA,length(daily_prices$base.HU[1:942])-length(data.frame(fcast2)[,3])),data.frame(fcast2)[,3]),
              Lo95 =c(rep(NA,length(daily_prices$base.HU[1:942])-length(data.frame(fcast2)[,4])),data.frame(fcast2)[,4]),
              Hi95 =c(rep(NA,length(daily_prices$base.HU[1:942])-length(data.frame(fcast2)[,5])),data.frame(fcast2)[,5]))
 
 ggplot(df, aes(time, real, group = 1)) +
   geom_line() +
   geom_line(aes(time, pred, group = 1), color = "blue")+
   geom_ribbon(aes(time, ymin = Lo95, ymax = Hi95), fill = "gray", alpha = 0.25) +
   geom_ribbon(aes(time, ymin = Lo80, ymax = Hi80), fill = "red", alpha = 0.25) +
   theme_light()
 
 
 ggplot(df2, aes(time, real, group = 1)) +
   geom_line() +
   geom_line(aes(time, pred, group = 1), color = "blue")+
   geom_ribbon(aes(time, ymin = Lo95, ymax = Hi95), fill = "gray", alpha = 0.25) +
   geom_ribbon(aes(time, ymin = Lo80, ymax = Hi80), fill = "red", alpha = 0.25) +
   theme_light()
```
## This is only when we have full 1095 data set forecasting
This is for the particular example
```{r}
 library(zoo)
 df <-
   data.frame(real = daily_prices$base.HU,
              pred = c(rep(NA,length(daily_prices$base.HU)-length(data.frame(fcast1)[,1])),data.frame(fcast1)[,1]),
              time =  zoo::as.Date(seq(as.Date("2017/1/1"), as.Date("2019/12/31"), by = "day"), format = "%Y-%m-%d"),
              Lo80 =c(rep(NA,length(daily_prices$base.HU)-length(data.frame(fcast1)[,2])),data.frame(fcast1)[,2]),
              Hi80 =c(rep(NA,length(daily_prices$base.HU)-length(data.frame(fcast1)[,3])),data.frame(fcast1)[,3]),
              Lo95 =c(rep(NA,length(daily_prices$base.HU)-length(data.frame(fcast1)[,4])),data.frame(fcast1)[,4]),
              Hi95 =c(rep(NA,length(daily_prices$base.HU)-length(data.frame(fcast1)[,5])),data.frame(fcast1)[,5])) 
 
 df2 <-
   data.frame(real = daily_prices$base.HU,
              pred = c(rep(NA,length(daily_prices$base.HU)-length(data.frame(fcast2)[,1])),data.frame(fcast2)[,1]),
              time =  zoo::as.Date(seq(as.Date("2017/1/1"), as.Date("2019/12/31"), by = "day"), format = "%Y-%m-%d"),
              Lo80 =c(rep(NA,length(daily_prices$base.HU)-length(data.frame(fcast2)[,2])),data.frame(fcast2)[,2]),
              Hi80 =c(rep(NA,length(daily_prices$base.HU)-length(data.frame(fcast2)[,3])),data.frame(fcast2)[,3]),
              Lo95 =c(rep(NA,length(daily_prices$base.HU)-length(data.frame(fcast2)[,4])),data.frame(fcast2)[,4]),
              Hi95 =c(rep(NA,length(daily_prices$base.HU)-length(data.frame(fcast2)[,5])),data.frame(fcast2)[,5])) 
 
 ggplot(df, aes(time, real, group = 1)) +
   geom_line() +
   geom_line(aes(time, pred, group = 1), color = "blue")+
   geom_ribbon(aes(time, ymin = Lo95, ymax = Hi95), fill = "gray", alpha = 0.25) +
   geom_ribbon(aes(time, ymin = Lo80, ymax = Hi80), fill = "red", alpha = 0.25) +
   theme_light()
 
 
 ggplot(df2, aes(time, real, group = 1)) +
   geom_line() +
   geom_line(aes(time, pred, group = 1), color = "blue")+
   geom_ribbon(aes(time, ymin = Lo95, ymax = Hi95), fill = "gray", alpha = 0.25) +
   geom_ribbon(aes(time, ymin = Lo80, ymax = Hi80), fill = "red", alpha = 0.25) +
   theme_light()
```


