---
title: "Master Thesis 5.5 GBM"
author: "Marijan Rancic"
date: "10/10/2020"
output: html_document
editor_options: 
chunk_output_type: inline
---

## IMPORTANT NOTE
This file is integral part of the PredictiveModeling_Master and will work only if data has been imported.
The purpose of this exercise is to evaluate best configuration for GBM.

## Load libraries
They should be already imported in the mail file.
```{r}
library(lubridate)
library(tidyverse)
library(gbm)
```

## Generate lagged values
We allow our models to include lagged values of the dependent variable (24 hours ago), as well as lagged values of predictors (24 hours ago and 1 hour ago) in addition to variables reflecting weekday, month, hour and the values of predictors in the hour for which the prediction is made.
This has been done in the PredictiveModeling_Master .

## Time series cross-validation
We use time series crossvalidation: take all data up to day t (thousands of observations) and predict for day t+1 (24 observations). But one testing sample is not enough so we do it for t from Jan 1 2019 to December 31 2019 - a total of 365 tests. this is an extremely computationally intensive approach, but it gives us a robust idea of what to expect from one-day ahead forecasts (what are typical mistakes, what is the average daily in 90% of cases? 99% of cases?).

Create a vector of 30 random test dates (parameter size=30) from 2019 data (a potential pool of 365 days). If you want to apply each model to a single date, just specify only this date as both start and end date and size=1 it will choose only from 1 date. In this example, the forecast will be for May 25 2019 only.
```{r}
library(lubridate)
set.seed(100)
test_dates<-sample(seq.Date(ymd("2019-01-25"),ymd("2019-10-25"),by=1),size=10)
```



----------------------------------------------------------------------------

## GBM for HU
```{r}
# set counter to 0 (will change in a loop)
k=0

# create empty vectors for MAE, MAPE and RMSE
gbm_mae_HU=numeric()
gbm_mape_HU=numeric()
gbm_rmse_HU=numeric()

# loop
for (i in seq_along(test_dates)){
date=test_dates[i]


# split sample
training<-subset(prices, ymd(datum)<date)
testing<-subset(prices, ymd(datum)==date)

# build model
gbm_HU <- gbm(HU~.-RS-HU_RS-`FlowHU-RS`-`FlowHU-RO`-hour_num,data=select_if(na.omit(training),is.numeric),
              n.trees=2000, 
              shrinkage = 0.05, 
              distribution = "tdist",
              interaction.depth = 1,
              cv.folds = 5)
forecast_gbm_HU <- predict.gbm(gbm_HU, na.omit(testing))
k=k+1
gbm_mae_HU[k]=mean(abs(forecast_gbm_HU-testing$HU))
gbm_mape_HU[k]=mean(abs((forecast_gbm_HU-testing$HU)/(testing$HU)*100))
gbm_rmse_HU[k]=sqrt(mean(abs(forecast_gbm_HU-testing$HU)^2))
}

## Print mean error measures based on all dates included in testing
mean(gbm_mae_HU)
mean(gbm_mape_HU)
mean(gbm_rmse_HU)
```

## Some settings for GBM
```{r}
gbm.perf(gbm_HU, method = "cv")
```


For the HU model
```{r}
summary(gbm_HU,plotit = FALSE)
```

F